---
title: "Scraping Race Results from the Web"
author: "Nicholas Melon & Spencer Rudnick"
date: "January 23, 2017"
output: html_document
---

Today we are working through Section 2.7 of the [race results chapter](http://moodle.cornellcollege.edu/pluginfile.php/101067/mod_resource/content/1/Runners.pdf).

#### Exercise 1
You wrote a function **extractResTable()** that takes as input the URL for a Web page containing race results, finds the preformatted text, and returns a character vector of the table of results.  Be sure to include code to handle the special cases of the 2000 and 2009 data.
```{r}
library(XML)

extractResTable =
       # Retrieve data from web site, find preformatted text,
       # return as a character vector.
  function(url)
  {
    doc <- htmlParse(url)
    preNode <- getNodeSet(doc, "//pre")
    txt <- xmlValue(preNode[[1]])
    els <- strsplit(txt, "\r\n")[[1]]
    return(els)
  }

ubase <- "http://www.cherryblossom.org/"
menURLs =
  c("cb99m.htm", "cb003m.htm", "results/2001/oof_m.html",
    "results/2002/oofm.htm", "results/2003/CB03-M.HTM",
    "results/2004/men.htm", "results/2005/CB05-M.htm",
    "results/2006/men.htm", "results/2007/men.htm",
    "results/2008/men.htm", "results/2009/09cucb-M.htm",
    "results/2010/2010cucb10m-m.htm",
    "results/2011/2011cucb10m-m.htm",
    "results/2012/2012cucb10m-m.htm")
urls = paste(ubase, menURLs, sep = "")

menTables <- lapply(urls, extractResTable)



# put your code here!
```

#### Exercise 2
Check your work by evaluating the following chunk of code.  Make sure that your output for this chunk matches the output in `cherry7_out.html`. 
```{r eval=FALSE}
ubase <- "http://www.cherryblossom.org/"
menURLs <-
  c("cb99m.htm", "cb003m.htm", "results/2001/oof_m.html",
    "results/2002/oofm.htm", "results/2003/CB03-M.HTM",
    "results/2004/men.htm", "results/2005/CB05-M.htm",
    "results/2006/men.htm", "results/2007/men.htm",
    "results/2008/men.htm", "results/2009/09cucb-M.htm",
    "results/2010/2010cucb10m-m.htm",
    "results/2011/2011cucb10m-m.htm",
    "results/2012/2012cucb10m-m.htm")
urls <- paste(ubase, menURLs, sep = "")
years <- 1999:2012
menTables <- mapply(extractResTable, url = urls, year = years)
names(menTables) <- years
sapply(menTables, length)
```