
\chapter{Cameron Seebach}

\begin{enumerate}
  \item Most programming languages require the use of brackets to
    enclose the index in a reference to an element of an array.
  \begin{enumerate}
    \item Identify a language the requires the use of parentheses
      to enclose the index in a reference to an element of an array.
    \item Why did the designers of the language choose parentheses
      rather than brackets?
    \end{enumerate}

  \begin{answer}

  \begin{enumerate}
    \item The original FORTRAN used parentheses to to enclose an array
      reference.
    \item At the time FORTRAN was first introduced, keypunch machines did not
      include the square bracket keys.
    \end{enumerate}

    \end{answer}

  \item What is the relationship between a lexeme and a token?

  \begin{answer}
    Lexemes and tokens are both produced by the lexer. A lexeme is a string that
    makes up some symbol in the language, while a token describes a whole
    category of lexemes.

    For example, an identifier lexeme is commonly a string of characters a-z and
    underscore. Regardless of what characters are chosen, the token is still
    identifier.
    \end{answer}

  \item
  \begin{enumerate}
    \item What kind of symbols are found at the internal nodes of a
      parse tree?
    \item What kind of symbols are found at the leaves of a parse tree?
    \end{enumerate}

  \begin{answer}

  \begin{enumerate}
    \item Non-terminal symbols are found at the internal nodes of a parse tree.
    \item Terminal symbols are found at the leaves of a parse tree. All the
      non-terminals have been replaced at the leaves of the tree.
    \end{enumerate}

    \end{answer}


  \item One of the most significant contributions from the developers
    of ALGOL 60 also limited the success of that language. What was
    that contribution?

  \begin{answer}

    Backus-Naur form described the syntax of ALGOL 60 in a way that was new
    and unfamiliar at the time.

    \end{answer}

  \item What problem were the creators of Common LISP trying to solve?

  \begin{answer}

    The creators of Common LISP sought to combine several different dialects of
    Lisp, including Scheme, into one language.

    \end{answer}

  \item What is an ambiguous context free grammar?

  \begin{answer}

    It is a context free grammar that can describe more than one distinct parse
    tree.

    \end{answer}

  \item Contrast the complexity of algorithms that can parse strings
    that conform to the most general kinds of context free grammars
    and the complexity of the algorithms that can parse strings that
    conform to the grammars of programming languages?

  \begin{answer}

    Algorithms that parse the most general CFGs have a time complexity of
    O(n**3). For a programming language grammar, algorithms exist that run in
    O(n) time.

    \end{answer}

  \item Java represents characters with Unicode. It is the first
    widely used programming language with this feature. What is the
    significance of this feature?

  \begin{answer}

    It is easier to write programs intended for an international audience in
    Java. Less special work is required to support character sets outside of
    ASCII.

    \end{answer}

  \item How does the binary coded decimal type differ from the
    floating point type?

  \begin{answer}

    Binary coded decimal typically allows for a fixed number of digits before
    the decimal point, and a fixed number of digits afterwards. By contrast,
    floating point types have a number of digits of precision which is not
    related to the location of the decimal point.

    Also, binary coded decimal is typically made up of groups of 4 bits (0-F),
    each four bits representing a digit (0-9). As a result, a binary coded
    decimal number takes more bits than a floating point number to store the
    same number.

    \end{answer}

  \item Identify a user-defined ordinal type in the Java programming
    language.

  \begin{answer}

    The user-defined ordinal type in Java is the Enum.

    \end{answer}

  \item Mathematicians and programmers might have different ideas
    about the precedence of Boolean operators. Explain.

  \begin{answer}

    In mathematics, AND and OR have the same precendence, but in the C language
    AND has a higher precedence.

    \end{answer}

  \item Programmers should use \verb+===+ rather than \verb+==+ to
    test the equality of the values of two expressions in JavaScript. Why?

  \begin{answer}

    Equality of strings and other objects is the best reason to use one over the
    other. \verb+===+ compares the content of two objects or strings,
    while \verb+==+ compares only the identity.

    \end{answer}

  \item Describe a hazard of allowing short-circuited evaluation
    of expressions and side effects in expressions at the same time.

  \begin{answer}

    If the expression with a side effect is placed second in a short-circuit
    operation, it may not always be applied if the operation short-circuits.

    This can lead to a situation where a programmer thinks that a mutating
    change has occurred, when in fact it has not.

    \end{answer}

  \item Briefly describe the three steps in the mark-sweep algorithm
    for garbage collection.

  \begin{answer}

    \begin{enumerate}
      \item Traverse the tree of objects reachable from the `root set', marking
        each as in-use.
      \item Scan all memory, freeing objects which are not in-use.
      \item Reset the in-use flag on each object still extant.

      \end{enumerate}

    \end{answer}

  \item What led Yukihiro Matsumoto to create the Ruby programming language?

  \begin{answer}

    Matsumoto wanted an easy, truly object-oriented scripting language, but
    didn't like the options available at that time in 1993.

    \end{answer}

  \item What did Microsoft aim to achieve with its development of the
    C\# language?

  \begin{answer}

    Microsoft aimed to have a flagship language similar to C and C++ for use
    with its new Common Language Runtime. Portability and general-purposeness
    were other important design goals.

    \end{answer}

  \end{enumerate}

\section{More questions for discussion and review.}

\begin{enumerate}
  \item The design of which machine influenced the design
    of the control statements in FORTRAN?

  \item How many different kinds of control statements
    must the designer of a programming language include
    in a language?

  \item What is the one question that applies in the
    design of all statements that allow selection or
    iteration?

  \item What is an advantage of requiring that
    the \textbf{then} and \textbf{else} clauses of
    an \textbf{if} statement be compound statements?

  \item How does the \textbf{switch} statement in C\#
    differ from the \textbf{switch} statement in Java?

  \item Distinguish between 2 statements in Ruby
    that correspond to Java's \textbf{switch} statement.

  \item Features of a programming language sometimes persist
    longer than a feature of computing hardware that inspired
    and supported that part of the language's design.
    Similarly, features of hardware sometimes persist longer
    than some parts of a language's design that were created
    to take advantage of that feature in hardware.

    Give examples.

  \item Who most famously warned of the dangers of using the
    \textbf{goto} statement? What did Donald Knuth have to
    say about the use of the \textbf{goto} statement?

  \item Describes Ada's \textbf{for} loop. Are there some
    kinds of iteration that might be easier in Ada than
    in Java? Easier in Java than in Ada?

  \item What does it mean to say that the guarded commands
    of Ada are non-deterministic?

  \item The header files in a C program contain function
    prototypes. What is a function prototype?

  \item Every method in a Ruby program belongs to a class.
    A programmer can place a definition of a method inside
    the definition of a class or outside of the definition
    of any class that the programmer writes. To which class
    does the method belong in the second case?

  \item Distinguish between positional and keyword parameters.

  \item Ruby blocks are closures. What does that mean?

  \item What is a pure function?

  \item Some languages give programmers means to define
    both functions and procedures. Java doe not. Is that
    a serious limitation?

  \item Declarations of formal parameters in an Ada procedure
    can include, in addition to the names and types of the
    parameters, reserved words that do not appear in declarations
    in Java programs. 
    What is the purpose of those reserved words?
 
  \item The C language imposes a constraint upon programmers
    who want to pass a multidimensional array to a function.
    What is the constraint? How did the design of the Java
    programming language eliminate that constraint for 
    programmers who use that language?

  \item An activation record contains a return
    address, a dynamic link, parameters, and
    local variables.
  \begin{enumerate}
    \item To what does the return address point?
    \item To what does the dynamic link point?
    \end{enumerate}

  \item The stack will contain multiple activation
    records for a single subprogram under what
    circumstances?

  \item How (or why?) does the LIFO protocol apply to
    calls to and returns from subprograms?

  \item Which important development in computer architecture
    has changed the way that the stack is used in some
    systems for facilitating calls to and returns from
    subprograms?

  \item A dynamic chain contains a history of what?

  \item Which two numbers are needed to compute
    the address of a local variable in a subprogram?

  \item How does a Ruby module differ from a class?

  \item Memory for variables can be allocated on the heap
    and on the stack. In which place or places is memory
    allocated for objects in C++? in Java?

  \item What problems were solved by the addition
    of genericity to Java?

  \item What is the purpose of the static chain?

  \item What is a singleton?

  \item What are the two parts of the definition 
    of an abstract data type?

  \end{enumerate}


