
\chapter{Huong Nguyen}

\begin{enumerate}
  \item Most programming languages require the use of brackets to
    enclose the index in a reference to an element of an array.
  \begin{enumerate}
    \item Identify a language the requires the use of parentheses
      to enclose the index in a reference to an element of an array.
    \item Why did the designers of the language choose parentheses
      rather than brackets?
    \end{enumerate}

  \begin{answer}

  \begin{enumerate}
    \item The languages are pre-90 Fortrans and PL/I
    \item Because there were no other suitable characters available at the time. Card punches did not include bracket characters.
    \end{enumerate}

    \end{answer}
    
  \item What is the relationship between a lexeme and a token?

  \begin{answer}

    Lexemes are the lowest-level syntactic units. A token of a language is a category of its lexemes.

    \end{answer}

  \item
  \begin{enumerate}
    \item What kind of symbols are found at the internal nodes of a
      parse tree?
    \item What kind of symbols are found at the leaves of a parse tree?
    \end{enumerate}

  \begin{answer}

  \begin{enumerate}
    \item Nonterminal symbols are found at the internal nodes of a parse trees.
    \item Terminal symbols are found at the leaves of a parse trees.
    \end{enumerate}

    \end{answer}


  \item One of the most significant contributions from the developers
    of ALGOL 60 also limited the success of that language. What was
    that contribution?

  \begin{answer}

    It was the use of BNF as the formal means of describing syntax. In 1960 it seemed strange and complicated.

    \end{answer}

  \item What problem were the creators of Common LISP trying to solve?

  \begin{answer}

    During the 1970s and early 1980s, a large number of different dialects of LISP were developed and used. This led to the familiar problem of lack of portability among programs written in the various dialects. Common LIPS was created to rectify the situation by combining the features of several dialects of LISP, including Scheme, into a single language.

    \end{answer}

  \item What is an ambiguous context free grammar?

  \begin{answer}

    An ambiguous grammar is a grammar that generates a sentential form for which there are two or more distinct parse trees. Context-free grammars were developed by Noam Chomsky as a class of natural languages, but became the primary method through which the syntax of programming languages are described. 

    \end{answer}

  \item Contrast the complexity of algorithms that can parse strings
    that conform to the most general kinds of context free grammars
    and the complexity of the algorithms that can parse strings that
    conform to the grammars of programming languages?

  \begin{answer}

Algorithms that can parse general kinds of context-free grammars have a cubic time complexity. But the grammars of programming languages are far less general and only a subset of general context-free grammars. Algorithms to parse them have O(n), or linear time complexity, which is more efficient.    

    \end{answer}

  \item Java represents characters with Unicode. It is the first
    widely used programming language with this feature. What is the
    significance of this feature?

  \begin{answer}

    Unicode includes the characters from most of the world's natural languages. Before Unicode was created, the 8-bit code ASCII was used to code character data and became inadequate for communication with computers around the world.

    \end{answer}

  \item How does the binary coded decimal type differ from the
    floating point type?

  \begin{answer}

    Floating-points store only approximations for many real values, while decimals store precise decimal values up to a specific range. Floating-point values are represented as fractions and exponents, typically taking 32 bits. Binary-coded decimals store one to two digits per byte, since a digit takes at least 4 bits, so they take up a lot of storage. 

    \end{answer}

  \item Identify a user-defined ordinal type in the Java programming
    language.

  \begin{answer}

    The java user-defined ordinal type is Enum. All enumeration types in Java are implicitly subclasses of the predefined class Enum.

    \end{answer}

  \item Mathematicians and programmers might have different ideas
    about the precedence of Boolean operators. Explain.

  \begin{answer}

    In the mathematics of Boolean algebras, the OR and AND operators must have equal precedence. However, the C-base languages assign a higher precedence to AND than OR. 

    \end{answer}

  \item Programmers should use \verb+===+ rather than \verb+==+ to
    test the equality of the values of two expressions in JavaScript. Why?

  \begin{answer}

    \verb+===+ prevents Javascript operands from being coerced, and in effect checks for type equivalence. For example, ``7'' \verb+==+ 7 evaluates to true in Javascript, while ``7'' \verb+===+ 7 evaluates to false.

    \end{answer}

  \item Describe a hazard of allowing short-circuited evaluation
    of expressions and side effects in expressions at the same time.

  \begin{answer}

    Allowing both at the same time may lead to errors where the side effect is part of the second boolean expression, and is not evaluated due to the expression having been short-circuited after the first.

    \end{answer}

  \item Briefly describe the three steps in the mark-sweep algorithm
    for garbage collection.

  \begin{answer}

    First, all cells in the heap are marked as garbage. Second, every pointer is traced into the heap, and mark all cells they reach as not-garbage. Finally, all cells still marked as garbage are returned to the list of available space.

    \end{answer}

  \item What led Yukihiro Matsumoto to create the Ruby programming language?

  \begin{answer}

    Matsumoto was dissatisfied with Perl and Python, which were not pure object-oriented languages.

    \end{answer}

  \item What did Microsoft aim to achieve with its development of the
    C\# language?

  \begin{answer}

    As the flagship language of .NET and Microsoft, C\# was meant to allow .NET to combine components written in all other .NET languages, such as Visual Basic and C++.

    \end{answer}

  \end{enumerate}



\section{More questions for discussion and review.}

\begin{enumerate}
  \item The design of which machine influenced the design
    of the control statements in FORTRAN?

  \item How many different kinds of control statements
    must the designer of a programming language include
    in a language?

  \item What is the one question that applies in the
    design of all statements that allow selection or
    iteration?

  \item What is an advantage of requiring that
    the \textbf{then} and \textbf{else} clauses of
    an \textbf{if} statement be compound statements?

  \item How does the \textbf{switch} statement in C\#
    differ from the \textbf{switch} statement in Java?

  \item Distinguish between 2 statements in Ruby
    that correspond to Java's \textbf{switch} statement.

  \item Features of a programming language sometimes persist
    longer than a feature of computing hardware that inspired
    and supported that part of the language's design.
    Similarly, features of hardware sometimes persist longer
    than some parts of a language's design that were created
    to take advantage of that feature in hardware.

    Give examples.

  \item Who most famously warned of the dangers of using the
    \textbf{goto} statement? What did Donald Knuth have to
    say about the use of the \textbf{goto} statement?

  \item Describes Ada's \textbf{for} loop. Are there some
    kinds of iteration that might be easier in Ada than
    in Java? Easier in Java than in Ada?

  \item What does it mean to say that the guarded commands
    of Ada are non-deterministic?

  \item The header files in a C program contain function
    prototypes. What is a function prototype?

  \item Every method in a Ruby program belongs to a class.
    A programmer can place a definition of a method inside
    the definition of a class or outside of the definition
    of any class that the programmer writes. To which class
    does the method belong in the second case?

  \item Distinguish between positional and keyword parameters.

  \item Ruby blocks are closures. What does that mean?

  \item What is a pure function?

  \item Some languages give programmers means to define
    both functions and procedures. Java doe not. Is that
    a serious limitation?

  \item Declarations of formal parameters in an Ada procedure
    can include, in addition to the names and types of the
    parameters, reserved words that do not appear in declarations
    in Java programs. 
    What is the purpose of those reserved words?
 
  \item The C language imposes a constraint upon programmers
    who want to pass a multidimensional array to a function.
    What is the constraint? How did the design of the Java
    programming language eliminate that constraint for 
    programmers who use that language?

  \item An activation record contains a return
    address, a dynamic link, parameters, and
    local variables.
  \begin{enumerate}
    \item To what does the return address point?
    \item To what does the dynamic link point?
    \end{enumerate}

  \item The stack will contain multiple activation
    records for a single subprogram under what
    circumstances?

  \item How (or why?) does the LIFO protocol apply to
    calls to and returns from subprograms?

  \item Which important development in computer architecture
    has changed the way that the stack is used in some
    systems for facilitating calls to and returns from
    subprograms?

  \item A dynamic chain contains a history of what?

  \item Which two numbers are needed to compute
    the address of a local variable in a subprogram?

  \item How does a Ruby module differ from a class?

  \item Memory for variables can be allocated on the heap
    and on the stack. In which place or places is memory
    allocated for objects in C++? in Java?

  \item What problems were solved by the addition
    of genericity to Java?

  \item What is the purpose of the static chain?

  \item What is a singleton?

  \item What are the two parts of the definition 
    of an abstract data type?

  \end{enumerate}


