
\chapter{Sam Caldwell}

\begin{enumerate}
  \item Most programming languages require the use of brackets to
    enclose the index in a reference to an element of an array.
  \begin{enumerate}
    \item Identify a language the requires the use of parentheses
      to enclose the index in a reference to an element of an array.
    \item Why did the designers of the language choose parentheses
      rather than brackets?
    \end{enumerate}

  \begin{answer}

  \begin{enumerate}
    \item Ada
    \item The designers of Ada specifically chose parentheses to
      enclose subscripts so there would be uniformity between array
      references and function calls in expressions, in spite of
      potential readability problems
    \end{enumerate}

    \end{answer}
    
  \item What is the relationship between a lexeme and a token?

  \begin{answer}

    A token represents functional groups of lexemes, which are
    representations of the lowest level of syntactic units

    \end{answer}

  \item
  \begin{enumerate}
    \item What kind of symbols are found at the internal nodes of a
      parse tree?
    \item What kind of symbols are found at the leaves of a parse tree?
    \end{enumerate}

  \begin{answer}

  \begin{enumerate}
    \item internal nodes have non-terminal symbols
    \item leaf nodes have terminal symbols
    \end{enumerate}

    \end{answer}


  \item One of the most significant contributions from the developers
    of ALGOL 60 also limited the success of that language. What was
    that contribution?

  \begin{answer}

    "Ironically, one of the most important contributions to computer
    science associated with ALGOL 60, BNF, was also a factor in its
    lack of acceptance. Although BNF is now considered a simple and
    elegant means of syntax descrip- tion, in 1960 it seemed strange
    and complicated."  pg. 57 (10th edition)

    \end{answer}

  \item What problem were the creators of Common LISP trying to solve?

  \begin{answer}

    The developers of Common LISP were trying to solve issues with
    portability among programs written in various dialects.

    \end{answer}

  \item What is an ambiguous context free grammar?

  \begin{answer}

    "A grammar that generates a sentential form for which there are
    two or more distinct parse trees is said to be ambiguous"

    \end{answer}

  \item Contrast the complexity of algorithms that can parse strings
    that conform to the most general kinds of context free grammars
    and the complexity of the algorithms that can parse strings that
    conform to the grammars of programming languages?

  \begin{answer}

    Parsing algorithms for unambiguous grammars are complicated and
    inneficient. In fact, " the complexity of such algorithms is
    $O(n^3)$"(pg. 180). On the other hand, the algorithms used for
    context free grammars are closer to the level of $O(n)$, which means
    the time they take is linearly related to the length of the string
    to be parsed. This is vastly more efficient than $O(n^3)$
    algorithms.

    \end{answer}

  \item Java represents characters with Unicode. It is the first
    widely used programming language with this feature. What is the
    significance of this feature?

  \begin{answer}

    The previously used ASCII was becoming obsolete with the
    globalization of business and the need for computers to
    communicate around the world. Java quickly becomes a global coding
    language based on it's acceptance of unicode - which includes
    (among other things) the Cryllic alphabet.

    \end{answer}

  \item How does the binary coded decimal type differ from the
    floating point type?

  \begin{answer}

    "Decimal types have the advantage of being able to precisely store
    dec- imal values, at least those within a restricted range, which
    cannot be done with floating-point"

    \end{answer}

  \item Identify a user-defined ordinal type in the Java programming
    language.

  \begin{answer}

    "There are two user-defined ordinal types that have been supported
    by programming languages: enumeration and subrange."

    \end{answer}

  \item Mathematicians and programmers might have different ideas
    about the precedence of Boolean operators. Explain.

  \begin{answer}

    Because in programming, arithmetic expressions can be the operands
    of relational expressions, and relational expressions can be the
    operands of Boolean expressions. For example the '=' sign - in
    mathmatics is signifies that two sides of an equation are the same
    value, wheras in programming it signifies the changing of value of
    a variable (or other things).

    \end{answer}

  \item Programmers should use \verb+===+ rather than \verb+==+ to
    test the equality of the values of two expressions in JavaScript. Why?

  \begin{answer}

    Because the '==' operator uses coercions to achieve equality,
    '===' is testing for it.

    \end{answer}

  \item Describe a hazard of allowing short-circuited evaluation
    of expressions and side effects in expressions at the same time.

  \begin{answer}

    "Suppose that short-circuit evaluation is used on an expression
    and part of the expres- sion that contains a side effect is not
    evaluated; then the side effect will occur only in complete
    evaluations of the whole expression."

    \end{answer}

  \item Briefly describe the three steps in the mark-sweep algorithm
    for garbage collection.

  \begin{answer}

    First, all cells in the heap have their indicators set to indicate
    they are garbage.  Second, Every pointer in the program is traced
    into the heap, and all reachable cells are marked as not being
    garbage Third, All cells in the heap that have not been
    specifically marked as still being used are returned to the list
    of available space
    \end{answer}

  \item What led Yukihiro Matsumoto to create the Ruby programming language?

  \begin{answer}

    "The motivation for Ruby was dissatisfaction of its designer with
    Perl and Python. Although both Perl and Python support
    object-oriented programming,14 nei- ther is a pure object-oriented
    language, at least in the sense that each has primi- tive
    (nonobject) types and each supports functions."(pg. 100)

    \end{answer}

  \item What did Microsoft aim to achieve with its development of the
    \verb+C#+ language?

  \begin{answer}

   "The purpose of \verb+C#+ is to provide a language for component-based
    software development, specifically for such development in the
    .NET Framework. In this environment, components from a variety of
    languages can be easily com- bined to form systems"

    \end{answer}

  \end{enumerate}



\section{More questions for discussion and review.}

\begin{enumerate}
  \item The design of which machine influenced the design
    of the control statements in FORTRAN?

  \item How many different kinds of control statements
    must the designer of a programming language include
    in a language?

  \item What is the one question that applies in the
    design of all statements that allow selection or
    iteration?

  \item What is an advantage of requiring that
    the \textbf{then} and \textbf{else} clauses of
    an \textbf{if} statement be compound statements?

  \item How does the \textbf{switch} statement in C\#
    differ from the \textbf{switch} statement in Java?

  \item Distinguish between 2 statements in Ruby
    that correspond to Java's \textbf{switch} statement.

  \item Features of a programming language sometimes persist
    longer than a feature of computing hardware that inspired
    and supported that part of the language's design.
    Similarly, features of hardware sometimes persist longer
    than some parts of a language's design that were created
    to take advantage of that feature in hardware.

    Give examples.

  \item Who most famously warned of the dangers of using the
    \textbf{goto} statement? What did Donald Knuth have to
    say about the use of the \textbf{goto} statement?

  \item Describes Ada's \textbf{for} loop. Are there some
    kinds of iteration that might be easier in Ada than
    in Java? Easier in Java than in Ada?

  \item What does it mean to say that the guarded commands
    of Ada are non-deterministic?

  \item The header files in a C program contain function
    prototypes. What is a function prototype?

  \item Every method in a Ruby program belongs to a class.
    A programmer can place a definition of a method inside
    the definition of a class or outside of the definition
    of any class that the programmer writes. To which class
    does the method belong in the second case?

  \item Distinguish between positional and keyword parameters.

  \item Ruby blocks are closures. What does that mean?

  \item What is a pure function?

  \item Some languages give programmers means to define
    both functions and procedures. Java doe not. Is that
    a serious limitation?

  \item Declarations of formal parameters in an Ada procedure
    can include, in addition to the names and types of the
    parameters, reserved words that do not appear in declarations
    in Java programs. 
    What is the purpose of those reserved words?
 
  \item The C language imposes a constraint upon programmers
    who want to pass a multidimensional array to a function.
    What is the constraint? How did the design of the Java
    programming language eliminate that constraint for 
    programmers who use that language?

  \item An activation record contains a return
    address, a dynamic link, parameters, and
    local variables.
  \begin{enumerate}
    \item To what does the return address point?
    \item To what does the dynamic link point?
    \end{enumerate}

  \item The stack will contain multiple activation
    records for a single subprogram under what
    circumstances?

  \item How (or why?) does the LIFO protocol apply to
    calls to and returns from subprograms?

  \item Which important development in computer architecture
    has changed the way that the stack is used in some
    systems for facilitating calls to and returns from
    subprograms?

  \item A dynamic chain contains a history of what?

  \item Which two numbers are needed to compute
    the address of a local variable in a subprogram?

  \item How does a Ruby module differ from a class?

  \item Memory for variables can be allocated on the heap
    and on the stack. In which place or places is memory
    allocated for objects in C++? in Java?

  \item What problems were solved by the addition
    of genericity to Java?

  \item What is the purpose of the static chain?

  \item What is a singleton?

  \item What are the two parts of the definition 
    of an abstract data type?

  \end{enumerate}


